### 2023-01-15 outputs/paper/paper.qmd
Update paper.qmd

---
title: "My document"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---